---
title: "HW2 Solutions"
author: "Alexander Cardazzi"
date: "Quarantine, 2020"
output:
  html_document:
    css: "format/buttonStyle.css"
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.align = "center")
```

**[All Materials Here](https://sites.google.com/view/alex-cardazzi/home/using-r)**

Use the sharks dataset for this HW.  This should be more slightly more difficult than the first homework.  You will likely have to do some googling, but I have hinted some functions to look into.  After each line, try to view the data to see if the results of your code match what you wanted it to do.

* Make a time series plot of the number of attacks per year.
* Make a frequency distribution for the ages of those being attacked.
* Plot the locations of the shark attacks.  Color the attacks by whether or not the attack was fatal.  Once this works, try and get fancy with the ```scales``` package.
* Make a new binary variable for whether the attack was on the west coast. (hint, ```?ifelse```)
* Make a binary variable for if the attack was unprovoked. (hint, ```?ifelse```)
* Make a binary variable for if there was "No injury" or a "minor" injury (hint, look for these inside the text of the Injury column.  hint, ```?grepl```)
* Make a binary variable for if the person attacked was male.  You may have to use both the Sex column and the Name column.
* Make summary stats for the sharks data.
* Use the binary variable for no injury or minor injury as your dependent variable.  Given an attack, what factors lead to more serious injuries?  Run a few regressions, have fun with it.
* Is there a variable that significantly changes the sample size of your regressions? Why
* Do you think these missing ages are [missing at random](https://www.ncbi.nlm.nih.gov/books/NBK493614/)?
* Are serious injuries more or less likely given missingness for the variable?

```{r, results = "hold"}
library("stargazer")
library("scales")
rm(list = ls())

sharks <- read.csv("C:/Users/alexc/Desktop/Empirical Workshop/data/sharks.csv", stringsAsFactors = FALSE)
#View(sharks)
```

###1 - Make a time series plot of the number of attacks per year.
```{r}
timeSeries <- table(sharks$Year)
plot(timeSeries, type = "l", ylab = "Frequency", xlab = "Year")
```

###2 - Make a frequency distribution for the ages of those being attacked.
```{r}
plot(table(as.numeric(sharks$Age)), ylab = "Frequncy", xlab = "Age")
```

###3 - Plot the locations of the shark attacks.  Color the attacks by whether or not the attack was fatal.
```{r}
sharks$fatalColor <- as.numeric(as.factor(sharks$Fatal..Y.N.))
#plot(sharks$Longitude, sharks$Latitude, pch = 19, col = c(1,2))
plot(sharks$Longitude, sharks$Latitude,
     col = alpha(sharks$fatalColor, .1), pch = 19)


```

###4 - Make a new binary variable for whether the attack was on the west coast.
```{r}
sharks$westCoast <- ifelse(sharks$Longitude < -100, 1, 0)
table(sharks$westCoast)
```

###5 - Make a binary variable for if the attack was unprovoked.
```{r}
table(sharks$Type)
sharks$unprovoked <- ifelse(sharks$Type == "Unprovoked", 1, 0)
table(sharks$unprovoked)
```

###6 - Make a binary variable for if there was "No injury" or a "minor" injury
```{r}
sharks$severe_injury <- ifelse(grepl("No injury", sharks$Injury) | grepl("minor", tolower(sharks$Injury)), 0, 1)
table(sharks$severe_injury)
```

###7 - Make a binary variable for if the person attacked was male.
```{r}
sharks$Sex <- ifelse(sharks$Name == "female", "F", ifelse(sharks$Name == "male", "M", sharks$Sex))
sharks$Sex <- ifelse(sharks$Sex == "", NA, sharks$Sex)
sharks$male <- ifelse(is.na(sharks$Sex), NA, ifelse(sharks$Sex == "M", 1, 0))
table(sharks$male)
```

###8 - Make summary stats for the sharks data.
```{r}
stargazer(sharks, type = "text")
```

###9 - Given an attack, what factors lead to more serious injuries?
```{r}
#exploring weather with these regressions
reg1 <- lm(severe_injury ~ westCoast, data = sharks)
reg2 <- lm(severe_injury ~ westCoast + Temperature + Precipitation + Wind.Speed + Visibility, data = sharks)
reg3 <- lm(severe_injury ~ westCoast*Temperature + Precipitation + Wind.Speed + Visibility, data = sharks)
stargazer(reg1, reg2, reg3,
          type = "text", omit.stat = c("ser", "f"))

#exploring the person/interaction
reg1 <- lm(severe_injury ~ westCoast*Temperature + unprovoked, data = sharks)
reg2 <- lm(severe_injury ~ westCoast*Temperature + male, data = sharks)
reg3 <- lm(severe_injury ~ westCoast*Temperature + unprovoked + male, data = sharks)
reg4 <- lm(severe_injury ~ westCoast*Temperature + unprovoked + male + as.numeric(Age), data = sharks)
reg5 <- lm(severe_injury ~ westCoast*Temperature + unprovoked + male + as.numeric(Age) + Year, data = sharks)
stargazer(reg1, reg2, reg3, reg4, reg5, 
          add.lines = list(c("Year FE", rep("N", 4), "Y")),
          type = "text", omit.stat = c("ser", "f"), omit = "Year")
```

###10 - Is there a variable that significantly changes the sample size of your regressions? Why
```{r}
#Including age decreases the the sample size by about 200.

table(sharks$Age[is.na(as.numeric(sharks$Age))])
#maybe we can start replacing teens with ~16, remove the "s" after some of these.  Still, 198 missing all together.
```

###12 - Are serious injuries more or less likely given missingness for the variable?
```{r}
lim <- is.na(as.numeric(sharks$Age))
t.test(sharks$severe_injury[lim], sharks$severe_injury[!lim])
t.test(sharks$westCoast[lim], sharks$westCoast[!lim])

#So it appears that missing ages are less likely to be severly injured and more likely to be on the east coast.  Perhaps this is spurious, but it might be a signal of something non-random.
```